Secure Composition of Robust and Optimising Compilers

### 2023-07-11 popl/main.tex
sec 4

\subsubsection{Monitor for \glsfirst{tms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

@@

\subsubsection{Monitor for \glsfirst{sms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{sms-uninteresting}{
% }{
%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% }{sms-uninteresting}
% %
% \typerule{sms-abort}{
% }{
%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% }{sms-abort}
  %
  \typerule{sms-use}{
    (\loc;\sandboxtag;m)\in\smsmonitor &
    n<m
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
  }{sms-use}
  %
  \typerule{sms-alloc}{
    (\loc;\sandboxtag;m)\notin\smsmonitor
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
  }{sms-alloc}
\end{center}
The state of the monitor for \gls{sms} is a set containing tuples of locations, control-tags, and the allocation size.
In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
The trace agreement is entirely straightforward and similar to the one for \gls{tms}.
% \begin{center}
%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
%   %
%   \typerule{sms-alloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
%   }{sms-alloc-authentic}
%   %
%   \typerule{sms-dealloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
%   }{sms-dealloc-authentic}
%   %
%   \typerule{sms-use-authentic}{
%   }{
%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
%   }{sms-use-authentic}
%   %
%   \typerule{sms-branch-authentic}{
%   }{
%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{sms-branch-authentic}
%   %
%   \typerule{sms-binop-authentic}{
%   }{
%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
%   }{sms-binop-authentic}
%   %
%   \typerule{sms-empty-authentic}{
%   }{
%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
%   }{sms-empty-authentic}
%   %
%   \typerule{sms-abort-authentic}{
%   }{
%     \smstraceagree[]{\lightning}{\absterm}
%   }{sms-abort-authentic}
% \end{center}

@@

The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
This is not intuitive in the sense that there are a number of examples in the security domain where sequential composition does not work~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.%
\MP{?canetti?} %% protocols under certain security notions are not secure when composed
The sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$ is defined as follows: given an $\src{L}$ program $\src{p}$, its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$.
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
\MPin{
	section 3.1 never defined the $\cap$ operator on properties, nor on classes.
	it uses the 'and'.
	rectify
}

@@

Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
Pretending that the following compilers are secure, consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
Were all these compilation steps robustly secure with respect to \gls{ms}, the resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\scriptsize\ttfamily]

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
Examples of this are present in industry:
Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
Both languages can be used at the same time in one project~\cite{androidstudio}.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
The definition of upper composition is as follows: given a program $\mathtt{p}$, its compiled $\trg{L}$ counterpart is obtained by plugging $\mathtt{p}$ into $\cc{\src{L}}{\trg{L}}$ if $\mathtt{p}\in\src{L}$ or by plugging $\mathtt{p}$ into $\cc{\obj{L}}{\trg{L}}$ if $\mathtt{p}\in\obj{L}$, written as $\cc{\src{L}+\obj{L}}{\trg{L}}$.
\MP{
	put into def env
}
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
The %proposed
definition of lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ is as follows: given a program $\src{p}$, its compiled $\trg{L}$ or $\obj{L}$ program is obtained by performing an internal decision and threading $\src{p}$ into $\cc{\src{L}}{\trg{L}}$ or into $\cc{\src{L}}{\obj{L}}$ depending on that decision.
\MP{into def}
Now, given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

>>>>

\subsubsection{Monitor for \glsfirst{tms}}\label{subsubsec:tms}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

@@

\subsubsection{Monitor for \glsfirst{sms}}\label{subsubsec:sms}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{sms-uninteresting}{
% }{
%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% }{sms-uninteresting}
% %
% \typerule{sms-abort}{
% }{
%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% }{sms-abort}
  %
  \typerule{sms-use}{
    (\loc;\sandboxtag;m)\in\smsmonitor &
    n<m
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
  }{sms-use}
  %
  \typerule{sms-alloc}{
    (\loc;\sandboxtag;m)\notin\smsmonitor
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
  }{sms-alloc}
\end{center}
The state of the monitor for \gls{sms} is a set containing tuples of locations, control-tags, and the allocation size.
In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
The trace agreement is entirely straightforward and similar to the one for \gls{tms}.
% \begin{center}
%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
%   %
%   \typerule{sms-alloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
%   }{sms-alloc-authentic}
%   %
%   \typerule{sms-dealloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
%   }{sms-dealloc-authentic}
%   %
%   \typerule{sms-use-authentic}{
%   }{
%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
%   }{sms-use-authentic}
%   %
%   \typerule{sms-branch-authentic}{
%   }{
%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{sms-branch-authentic}
%   %
%   \typerule{sms-binop-authentic}{
%   }{
%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
%   }{sms-binop-authentic}
%   %
%   \typerule{sms-empty-authentic}{
%   }{
%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
%   }{sms-empty-authentic}
%   %
%   \typerule{sms-abort-authentic}{
%   }{
%     \smstraceagree[]{\lightning}{\absterm}
%   }{sms-abort-authentic}
% \end{center}

@@

The main result is that secure compilers in the robust compilation framework~\cite{abate2019jour} compose {\em sequentially}.
This is not intuitive in the sense that in the security domain, composition does not work without additional generalizations~\cite{canetti2006univcomp,mccullough2012compo,fabian2022automatic}.
The \bul{sequential composition of compilers $\cc{\src{L}}{\trg{L}}$ and $\cc{\trg{L}}{\obj{L}}$} is defined as follows:
Given an $\src{L}$ program $\src{p}$ and compilers $\cc{\src{L}}{\trg{L}},\cc{\trg{L}}{\obj{L}}$, \rul{its compiled $\obj{L}$ counterpart is obtained by plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$}.
\begin{definition}
  \bul{$\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$} $\isdef$ Given $\src{p}$, yield \rul{$\cc{\trg{L}}{\obj{L}}\left(\cc{\src{L}}{\trg{L}}(\src{p})\right)$}
\end{definition}

Consider the compilation chain for $\src{TypeScript}$.
First, $\src{TypeScript}$ programs are translated to $\trg{JavaScript}$ which, e.g., V8~\cite{googlev8} eventually compiles in parts to $\irl{Ignition BC}$.
The following theorem establishes what happens if all these compilation steps were robustly secure with respect to \gls{ms}:
The resulting $\irl{Ignition BC}$ code would be \gls{ms} regardless of the context the binary runs in.

Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\trg{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their sequential composition $\cc{\src{L}}{\trg{L}}\circ\cc{\trg{L}}{\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

Since the composition of secure compilers is again a secure compiler, the theorem generalises to a whole chain of $n$ secure compilers.
It is worth pointing out that \gls{ms} (\Cref{subsubsec:ms}) is an example of a practically relevant intersection of classes, namely, \gls{tms} and \gls{sms} (\Cref{subsubsec:tms,subsubsec:sms}).
However, the properties need to be lifted~\cite{clarkson2008hyper} to sets of properties by applying the powerset operation.

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left] (progunopt) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progunopt] (progoptcf) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]

@@

      \node[minimum width=3cm,minimum height=2.5cm,draw=black,very thick,align=left,right=1.0 of progoptcf] (progoptcfdce) {\begin{lstlisting}[language=ml,basicstyle=\small\ttfamily]

@@

\subsection{Secure Upper and Lower Composition}\label{sec:other-compos}
Besides sequential composition, there are two other compositions, namely an {\em upper}, i.e., a compiler that takes multiple inputs and yields one output, and a {\em lower} composition, i.e., a compiler that takes one input and yields multiple outputs.
\bul{Define the upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$} as follows:
Given a program \texttt{p}, its compiled counterpart is obtained by \rul{plugging \texttt{p} into $\cc{\src{L}}{\trg{L}}$ if $\texttt{p}\in\src{L}$} or by \oul{plugging \texttt{p} into $\cc{\obj{L}}{\trg{L}}$ if $\texttt{p}\in\obj{L}$}.
\begin{definition}[Upper Composition]
  \bul{$\cc{\src{L}+\obj{L}}{\trg{L}}$} $\isdef$ Given \texttt{p}, yield
  $\left\{\mbox{\begin{tabular}{c}
    {\rul{if $\texttt{p}\in\src{L}$, then $\cc{\src{L}}{\trg{L}}(\texttt{p})$}} \\
    \mbox{\oul{if $\texttt{p}\in\obj{L}$, then $\cc{\obj{L}}{\trg{L}}(\texttt{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

Examples of this are present in industry:
Consider the Java Virtual Machine bytecode $\trg{JVM BC}$, which is a popular target for programming language designers due to its high performance and relevance in industry.
Compilers for several programming languages have it as their target language, some popular instances are $\src{Java}$ and $\obj{Kotlin}$.
Technically speaking, they both compile to class files and $\obj{Kotlin}$ objects are considered to be the same as $\src{Java}$ objects at that point.
Both languages can be used at the same time in one project~\cite{androidstudio}.
A compiler that accepts both $\src{Java}$ and $\obj{Kotlin}$ code translating to the same target language or intermediate representation performs a kind of {\em upper} composition.
Now, the following theorem tells us what happens if these are secure:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\obj{L}}{\trg{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their upper composition $\cc{\src{L}+\obj{L}}{\trg{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.

@@

Dually, the {\em lower} composition is concerned about compilers that accept the same source but yield different target languages. %and the same theoretical results apply.
\bul{Define the lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$} as follows:
Given a program $\src{p}$, its compiled counterpart is obtained by \rul{plugging $\src{p}$ into $\cc{\src{L}}{\trg{L}}$} or by \oul{plugging \texttt{p} into $\cc{\src{L}}{\obj{L}}$}, respectively, \pul{based on the internal decision}.
\begin{definition}[Lower Composition]
  \bul{$\cc{\src{L}}{\trg{L}+\obj{L}}$} $\isdef$ Given $L$ and $\src{p}$, yield
  $\left\{\mbox{\begin{tabular}{c}
    {\pul{if $L=\trg{L}$, then} \rul{$\cc{\src{L}}{\trg{L}}(\src{p})$}} \\
    \mbox{\pul{if $L=\obj{L}$, then} \oul{$\cc{\src{L}}{\obj{L}}(\src{p})$}} \\
  \end{tabular}}\right.$
%
%
%  $\begin{cases}
%                                                                            \rul{\cc{\src{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\src{L}} \\
%                                                                            \cc{\obj{L}}{\trg{L}}(\texttt{p}) & \text{if }\texttt{p}\in\obj{L}
%                                                                          \end{cases}$
\end{definition}

Consider two compilers both accepting $\src{LLVM IR}$~\cite{lattner2004llvm} and one of them emits $\trg{x86\_64}$, while the other emits $\trg{ARMv8}$.
It is intuitive that they are in some sense composed in the LLVM framework, but the decision of when to use one over the other is inherently {\em internal} to the formalisation effort of this kind of composition.
For example, the user of this compiler provides an explicit flag that instructs to emit $\trg{x86\_64}$ or the framework itself detects the target platform via heuristics, such as supported instructions.
The following theorem demonstrates what happens if the involved compilres are secure:
Given \bul{$\cc{\src{L}}{\trg{L}}$ robustly preserves $\class[_{1}]$} and \rul{$\cc{\src{L}}{\obj{L}}$ robustly preserves $\class[_{2}]$}, it follows that \oul{their lower composition $\cc{\src{L}}{\trg{L}+\obj{L}}$ robustly preserves the intersection of classes $\class[_{1}]$ and $\class[_{2}]$}.
